This script is to collate output files from different methods into a unified results files

```{r}
library(tidyverse)
library(treeio)
library(ggtree)
library(rjson)

# Analysis-specific parameters
outbreak_annotation <- "CONF_A"
data_dir <- "nextflow/results/boston_confa"
full_metadata <- read.delim(paste0(data_dir, "/metadata_with_context.tsv"))
outbreak_metadata <- full_metadata %>% filter(division == outbreak_annotation)
```

# Load tabular results files
```{r}
clustertracker_results <- read.delim(paste0(data_dir, "/clustertracker/introductions.tsv"))
```

# Deal with tree results files
Convert timetree output to New Hampshire eXtended format
```{bash}
cd nextflow/results/boston_confa/mugration/
grep "^ Tree" annotated_tree.nexus | sed "s/ Tree tree1=//g" | sed "s/&/&&NHX:/g" > annotated_tree.nhx
cd ../../../../
```

Get clusters from treetime mugration output
```{r}
mugration_treeio <- treeio::read.nhx(paste0(data_dir, "/mugration/annotated_tree.nhx"))
node_annotations <- as.data.frame(as_tibble(mugration_treeio))
mugration_phylo <- ape::as.phylo(mugration_treeio)

# pre-order tree traversal to find outbreak MRCA nodes
mugration_phylo <- ape::reorder.phylo(mugration_phylo, order = "postorder", index.only = F)
postorder_edges <- mugration_phylo$edge

cluster_mrcas <- c()
mrca_found_yet <- F
for (i in nrow(postorder_edges):1) {
    child <- postorder_edges[i, 2]
    annotation <- node_annotations[child, 5]
    if (annotation == outbreak_annotation & !mrca_found_yet) {
        print(paste("Found an outbreak MRCA at node", child))
        cluster_mrcas <- c(cluster_mrcas, child)
        mrca_found_yet <- T
    } else if (annotation != outbreak_annotation) {
        mrca_found_yet <- F
    }
}

# deal with polytomies: if one MRCA is a descendent of another, consider only the node with more descendents
sub_mrcas <- c()
for (i in length(cluster_mrcas):2) {
    next_mrca <- cluster_mrcas[i - 1]
    descendents_of_next_mrca <- phytools::getDescendants(tree = mugration_phylo, node = next_mrca)
    if (cluster_mrcas[i] %in% descendents_of_next_mrca) {
        sub_mrcas <- c(sub_mrcas, cluster_mrcas[i])
    }
}
cluster_mrcas_depolytomied <- setdiff(cluster_mrcas, sub_mrcas)
```

Get mugration outbreak clusters is a tabular format
```{r}
is_first <- T
for (mrca in cluster_mrcas_depolytomied) {
    descendant_idxs <- phytools::getDescendants(tree = mugration_phylo, node = mrca)
    tip_idxs <- descendant_idxs[descendant_idxs <= length(mugration_phylo$tip.label)]
    tip_names <- mugration_phylo$tip.label[tip_idxs]
    res <- data.frame(
        sample = tip_names,
        introduction_node = paste(outbreak_annotation, "node", mrca, sep = "_"),
        cluster_size = length(tip_names),
        region = node_annotations[tip_idxs, 5])
    if (is_first) {
        mugration_results <- res
    } else {
        mugration_results <- rbind(mugration_results, res)
    }
}

# ggtree(mugration_treeio) + 
#     geom_label(aes(label = division)) +
#     geom_tiplab()
# ggsave("~/Downloads/tree.png", width = 20, height = 20)    
```

Parse HIV-TRACE json output
// TODO: these results only contain clusters > size 1, missing singletons
```{r}
clusters_json <- readr::read_file(paste0(data_dir, "/hiv_trace/clusters.json"))
clusters_list <- rjson::fromJSON(json_str = clusters_json)

cluster_idxs <- clusters_list$trace_results$Nodes$cluster$values
clustered_sample_names <- clusters_list$trace_results$Nodes$id

cluster_results <- data.frame(
    sample = clustered_sample_names,
    introduction_node = cluster_idxs)

outbreak_sample_names <- outbreak_metadata$strain
sample_names <- full_metadata$strain
n_clusters <- clusters_list$trace_results$`Network Summary`$Clusters
n_singletons <- clusters_list$trace_results$`Network Summary`$Singletons

singleton_results <- data.frame(
    sample = sample_names[!(sample_names %in% clustered_sample_names)],
    introduction_node = n_clusters:(n_clusters + n_singletons - 1))  # hivtrace clusters are zero-indexed

hivtrace_results <- rbind(cluster_results, singleton_results) %>% 
    group_by(introduction_node) %>% 
    mutate(cluster_size = n()) %>%
    mutate(region = case_when(
        sample %in% outbreak_sample_names ~ outbreak_annotation, T ~ "Other"))
```

# Combine all results
```{r}
full_results <- rbind(
    clustertracker_results %>% 
        select(sample, introduction_node, cluster_size, region) %>%
        mutate(method = "clustertracker"),
    mugration_results %>% mutate(method = "mugration"),
    hivtrace_results %>% mutate(method = "hivtrace")
)

# TODO: parse mugration results to be per-sample for all samples, not just those under cluster introduction nodes
full_results %>%
    group_by(method) %>%
    summarize(n_samples = n())

# inspect results
full_results %>%
    group_by(method, introduction_node) %>%
    summarize(
        n_outbreak_samples = sum(region == outbreak_annotation),
        n_total_samples = n()) %>%
    arrange(desc(n_outbreak_samples))

# save results
write.table(
    x = full_results, 
    file = paste0(data_dir, "/collated_results.txt"),
    row.names = F)
```